A Stopped Negative Binomial Distribution

### 2016-06-07 stat-and-prob-letters/snb-paper.tex
More of Dan's comments integrated.

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients 
enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the fourth outcome and on in the eleventh.

@@

In this work we derive the distribution of the number of enrollees needed
for either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB) and some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
The rest of this paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the Binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the posterior distribution using a Beta prior distribution
for $p$. Section 6 provides a brief discussion on the use of the distribution
in clinical trials along with future avenues for generalization.

@@

The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
The probability mass function is
\begin{equation} \label{eqn:pmf}
\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is the {\emph indicator function}, taking a value 
of one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability mass.

@@

The probability a given realization of $\mathbf{X}$ reaches $s$ at
the $k$th outcome is the probability that, at time $k-1$ there are $s-1$
successful outcomes and $k-s$ unsuccessful outcomes multiplied by
the probability of a success at time $k$. This expression is given
in (\ref{eqn:N}). 
Similarly, probability a given realization reaches $k-t$
is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
and $t-1$ unsuccessful outcomes multiplied by the probability of an
unsuccessful outcome at time $k$.

@@

To show that (\ref{eqn:pmf}) sums to one, define
\begin{align} \label{eqn:sum_proof}
R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{align}
where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
Then $R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

@@

Next, we consider an interim analysis of a clinical trial after $s'$ 
successes and $t'$ failures have been observed for $s' < s$ and $t' < t$.
\begin{corollary} \label{conditional_distribution}
The number of subsequent enrollements needed 
to reach an endpoint is SNB($p$, $s-s'$, $t-t'$).
\end{corollary}
Having observed $s'$
successes and $t'$ failures, there are $s-s'$ successes needed to reach the
success endpoint and $t-t'$ failures needed to reach the failure endpoint.

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters 
($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
$s$ successes before $t$ failures, grey indicates
mass contributed by reaching $t$ failures first. \label{shapes.fig}}
\end{figure}

@@

There is a close connection between the tail probabilities of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to the probability of at most $s-1$ successes in a Binomial distribution. 
That is, the probability that the number of successes is at least $s$ in the 
Binomial model is the same that the trial succeeds (reaches $s$) in the SNB 
model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [\text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{align*}
\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \mathcal{I}_{1-p}.
\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [\text{trial succeeds}] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [\text{trial succeeds}] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in (\ref{eqn:first_sum}) satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows we must restrict
$x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

\section{The Posterior Distribution}

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The posterior probability mass function of the Stopped Negative Binomial 
distribution with a 
Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose t-1} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
\end{align*}
The result in (\ref{eqn:posterior}) follows by the definition of the 
Beta function.
\end{proof}

To better understand the posterior distribution, let us once again return to 
our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
with shape parameters $\alpha = 0.045 c$ and 
$\beta = 0.955 c$ where $c$ will take values between one and 10. 
This parameterization allows us to examine the relationship between 
uncertainty in $p$ and the SNB.

\begin{figure}[t!]
\centering
\subfloat[The expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[The variance of the sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The expected size and variance of the posterior SNB varying the uncertainty 
in the prior distribution of $p$. A larger value of $c$
corresponds to a Beta prior with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}

Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
and variance of the posterior SNB. As the certainty increases the 
posterior values converge to corresponding expected values with 
$p = 0.045$. It may be noted that convergence is relatively slow. This
is because when $c$ is small, the prior is heavily skewed. As $c$ increases
this effect is minimized.

Slow convergence in the mean will occur when $p$ takes values close to 
zero or one. In these cases it may be preferable to select shape paramters
based on the median or mode of the beta distribution. These values
will vary less in the uncertainty of the distribution.

\section{Discussion}

The SNB can be used to extend clinical trial methodology
in at least two areas. First, it can be used to construct nested criteria
for early stopping and efficacy. Current approaches are often inspired by
the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
a small number of patients are enrolled for an initial stage. For patients,
success or failure is assessed. If at least a pre-specified number of
successes are observed then the trial enters a second stage with more
patients. If a sufficient number of successes are not observed then the
trial fails.

@@

A second avenue for methodological innovation comes directly from
Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
Our knowledge of the response probability $p$ comes from published reports
yielding a prior distribution.
While a trial is being conducted the response probabilities are 
updated as new results are received. The number of individual successes
and failure are used to learn the shape characteristics of the corresponding
Beta distribution and resulting posterior
distribution. An estimate of the probability of reaching the success (and
failure) endpoints can be provided at each step of the trial.

>>>>

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients 
enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the fourth outcome and on in the eleventh.
Since the success boundary is reached, we say that the trial succeeds,
or more correctly, the treatment succeeds.

@@

In this work we derive the distribution of the number of enrollees needed
to observe either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB). Some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
The rest of this paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution function
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the Binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the posterior distribution using a Beta prior distribution
for $p$. Section 6 provides a brief discussion on the use of the distribution
in clinical trials along with future avenues for generalization.

@@

The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
The probability mass function is
\begin{equation} \label{eqn:pmf}
\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is the {\emph indicator function}, taking the value 
of one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability mass.

@@

The probability a given realization of $\mathbf{X}$ reaches $s$ at
the $k$th outcome is the probability that, at time $k-1$, there are $s-1$
successful outcomes and $k-s$ unsuccessful outcomes multiplied by
the probability of the final success at time $k$. This expression is given
in (\ref{eqn:N}). 
Similarly, the probability a given realization reaches $k-t$
is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
and $t-1$ unsuccessful outcomes multiplied by the probability of a final
unsuccessful outcome at time $k$.

@@

To show that (\ref{eqn:pmf}) sums to one, define
%\begin{align} \label{eqn:sum_proof}
%R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
%  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
%\end{align}
\begin{equation*} 
R = \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t).
\end{equation*}
If we substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
$R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

@@

Next, we consider an interim analysis of a clinical trial after $s'$ 
successes and $t'$ failures have been observed for $s' < s$ and $t' < t$.
\begin{corollary} \label{conditional_distribution}
The number of subsequent enrollments needed 
to reach either endpoint behaves as SNB($p$, $s-s'$, $t-t'$).
\end{corollary}
Having observed $s'$
successes and $t'$ failures, there are $s-s'$ successes needed to reach the
success endpoint and $t-t'$ failures needed to reach the failure endpoint.

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters 
($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
$s$ successes before $t$ failures. Grey indicates
mass contributed by reaching $t$ failures first. \label{shapes.fig}}
\end{figure}

@@

There is a close connection between the tail probabilities of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to the probability of at most $s-1$ successes in a Binomial distribution. 
That is, the probability that the number of successes is at least $s$ in the 
Binomial model is the same that the trial succeeds (reaches $s$) in the SNB 
model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [\text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{equation*}
\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}.
\end{equation*}
%\begin{align*}
%\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
%  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
%  &= 1 - \mathcal{I}_{1-p}.
%\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [\text{trial succeeds}] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [\text{trial succeeds}] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in (\ref{eqn:first_sum}) satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows we must restrict
$x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

The SNB's ability to approximate the geometric, normal, and gamma 
distributions follow from it generalizing the negative binomial. To 
recover the MGF of the negative binomial consider the case where
$t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
beta function in the first term goes to one and zero in the second term
and we are left with the MGF of the negative binomial distribution. 
When $s=1$ this is the same as the geometric distribution. The negative
binomial can therefore be seen as a sum of i.i.d. geometric distributions.
For an appropriately large number of samples the central limit theorem
hold yielding a normal approximation.

Drawing a connection to the gamma distribution is more complicated.
However it is a well-studied problem in the literature (see 
\cite{Ord1968,Guenther1972,Best1974} for examples). Methods for these
approximation generally work by equating cumulants and then showing 
that errors are not too far off.

The lower-half normal distribution can be approximated by setting $s=t$
for appropriately large $s$ and $t$ and $p = 0.5$. In this
case the SNB can be viewed as identical, negative binomials
approximating a normal and truncated at the median.

\section{The Compound Distribution}

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The compound distribution of the Stopped Negative Binomial 
distribution with a Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose t-1} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
\end{align*}
The result in (\ref{eqn:posterior}) follows by the definition of the 
Beta function.
\end{proof}

To better understand the compound distribution, let us once again return to 
our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
with shape parameters $\alpha = 0.045 c$ and 
$\beta = 0.955 c$ for any $c > 0$.
This parameterization allows us to examine the relationship between 
uncertainty in $p$ and the SNB.

\begin{figure}[t!]
\centering
\subfloat[The expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[The variance of the sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The expected size and variance of the compound SNB varying the uncertainty 
in the prior distribution of $p$. A larger value of $c$
corresponds to a Beta prior with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}

Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
and variance of the compound SNB. As the certainty increases the 
compound distribution sample values converge to corresponding expected 
values with 
$p = 0.045$. It may be noted that convergence is relatively slow. This
is because when $c$ is small, the prior is heavily skewed. As $c$ increases
this effect is minimized.

Slow convergence in the mean will occur when $p$ takes values close to 
zero or one. In these cases it may be preferable to select shape parameters
based on the median or mode of the beta distribution. These values
will vary less in the uncertainty of the distribution.

%\section{Discussion}
%
%The SNB can be used to extend clinical trial methodology
%in at least two areas. First, it can be used to construct nested criteria
%for early stopping and efficacy. Current approaches are often inspired by
%the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
%a small number of patients are enrolled for an initial stage. For patients,
%success or failure is assessed. If at least a pre-specified number of
%successes are observed then the trial enters a second stage with more
%patients. If a sufficient number of successes are not observed then the
%trial fails.

@@

%A second avenue for methodological innovation comes directly from
%Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
%Our knowledge of the response probability $p$ comes from published reports
%yielding a prior distribution.
%While a trial is being conducted the response probabilities are 
%updated as new results are received. The number of individual successes
%and failure are used to learn the shape characteristics of the corresponding
%Beta distribution and resulting posterior
%distribution. An estimate of the probability of reaching the success (and
%failure) endpoints can be provided at each step of the trial.
